Surrogate model for aligned-spin binary neutron star waveforms using gaussian process regression

### 2017-07-20 paper/paper.tex
Include more explicit formulas in the EOB section

Let us now discuss the construction of the full waveform, from the inspiral to the post-merger stage. In this work we only consider the dominant mode, $(\ell,m)=(2,2)$. The inspiral-plunge GW emission is obtained by evaluating Eq.~(\ref{hlminspplunge}) along the orbital dynamics up to $t=t_{\textrm{match}}^{22}$, that is the time when either the amplitude or the GW frequency of $h_{22}^{\textrm{insp-plunge}}$ reaches a global maximum, whichever occurs first. We model the post-merger GW emission --- i.e., at $t>t_{\textrm{match}}^{22}$ --- by a signal that tapers down in amplitude and has a constant GW frequency. In particular, we employ the amplitude of the merger-ringdown of the BBH model (as given by the magnitude of Eq.~(2.10) of Ref.~\cite{Bohe:2016gbl}), using as inputs the mass ratio and spins of the BNS/NSBH of interest. This choice is in reasonable qualitative agreement with the post-merger behavior of the two equal-mass nonspinning NR BNS simulations that were analyzed in Ref.~\cite{Hinderer:2016eia}. The constant value of the GW frequency is taken to be that of $h_{22}^{\textrm{insp-plunge}}$ at time $t=t_{\textrm{match}}^{22}$. A more sophisticated, NR-informed, time-domain model of the post-merger emission will be part of future investigations.

@@

As described above, an EOB waveform with dynamical tides depends on 11 intrinsic parameters: the mass ratio $0\leq q=M_B/M_A\leq1$, the dimensionless spin components along the angular momentum of the binary $-1 \leq \chi_{A,B} \leq1$, the adiabatic quadrupolar tidal polarizabilities $\Lambda_2^{A,B}$, the adiabatic octupolar tidal polarizabilities $\Lambda_3^{A,B}$, the $\ell=2,3$ $f$-mode angular frequencies $\omega_{0\ell}^{A,B}$. To reduce the dimensionality of the intrinsic parameter space, we exploit universal (i.e., equation-of-state-independent) relations that are available in the literature. In particular, Eq.~(60) of Ref.~\cite{Yagi:2013sva} with the coefficients from the first row of Table I therein expresses $\Lambda_3^{A,B}$ as a function of $\Lambda_2^{A,B}$ with an accuracy to within about 10\%, depending on the equation of state and the value of $\Lambda_2^{A,B}$. Moreover, Eq. (3.5) of Ref.~\cite{Chan:2014kua} with the coefficients from the first column of Table I therein gives $\omega_{02}^{A,B}$ as a function of $\Lambda_2^{A,B}$ to within a few percent, while the same equation but with the coefficients from the second column of Table I therein gives $\omega_{03}^{A,B}$ as a function of $\Lambda_3^{A,B}$ to within a few percent. Thanks to these relations, the waveform only depends on 5 intrinsic parameters, which we denote as $\boldsymbol{x}=\{q, \chi_A,\chi_B, \Lambda_2^A,\Lambda_2^B\}$. We note that this list does not include the total mass $M$ of the system. Of course, the point-mass part of the model is scale-invariant. As to the inspiral-plunge phase, in this model tidal corrections only depend on $X_A=1/(1+q)$, $X_B=q/(1+q)$, and $\Lambda_2^{A,B}$, which is inherent to the post-Newtonian analysis of tidal effects~\cite{Flanagan:2007ix}. Finally, the simple model of the post-merger signal that we employ does not depend on $M$ either, although a more sophisticated description will have to include such dependence to account for the phenomenology that is observed in NR simulations of this phase.

>>>>

Let us now discuss the construction of the full waveform. In this work we only consider the dominant mode, $(\ell,m)=(2,2)$. The inspiral-plunge GW emission is obtained by evaluating Eq.~(\ref{hlminspplunge}) along the orbital dynamics up to $t=t_{\textrm{match}}^{22}$. Let $t_{\textrm{peak}}^{\textrm{amp}}$ be the time when $|h_{22}^{\textrm{insp-plunge}}|$ is maximum. Let $t_{\textrm{peak}}^{\textrm{freq}}$ be the time when the inspiral-plunge GW frequency $\omega_{22}^{\textrm{insp-plunge}}$ is maximum. Then we put
\begin{equation}
t_{\textrm{match}}^{22} = \min \left(t_{\textrm{peak}}^{\textrm{amp}},t_{\textrm{peak}}^{\textrm{freq}}\right)\,.
\end{equation}
We model the post-merger GW emission by a signal that tapers down in amplitude and has a constant GW frequency. In particular, we employ the amplitude of the merger-ringdown of the underlying BBH model~\cite{Bohe:2016gbl}, using as inputs the mass ratio and spins of the BNS/NSBH of interest. Explicitly,
\begin{equation}
|h_{22}^{\textrm{post-mrg}}(t)|=\nu \tilde{A}_{22}(t)\exp{\left[-(t-t^{22}_{\textrm{match}})/\tau_{220}\right]}\,,
\end{equation}
where $\nu=M_A M_B/M^2$ is the symmetric mass ratio, $\tilde{A}_{22}(t)$ is a monotonic and bounded function given in Eq.~(2.12) of Ref.~\cite{Bohe:2016gbl}, and $\tau_{220}$ is the decay time of the least-damped quasinormal mode of the remnant BH that forms after the BBH merger. This choice is in reasonable qualitative agreement with the post-merger behavior of the two equal-mass nonspinning NR BNS simulations that were analyzed in Ref.~\cite{Hinderer:2016eia}. The constant value of the GW frequency is taken to be that of $h_{22}^{\textrm{insp-plunge}}$ at time $t=t_{\textrm{match}}^{22}$, so that the post-merger phase evolution reads
\begin{align}
\phi_{22}^{\textrm{post-mrg}}(t)&=\omega_{22}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{22})(t-t_{\textrm{match}}^{22})\nonumber\\
&+\phi_{22}^{\textrm{inspiral-plunge}}(t_{\textrm{match}}^{22})\,
\end{align}
where $\phi_{22}^{\textrm{inspiral-plunge}}$ is the phase of $h_{22}^{\textrm{insp-plunge}}$. A more sophisticated, NR-informed, time-domain model of the post-merger emission will be part of future investigations. Finally, the full waveform reads
\begin{align}
h_{22}(t) &= h_{22}^{\textrm{insp-plunge}}(t)\, \theta\left(t_{\textrm{match}}^{22}-t\right)\nonumber\\
&+h_{22}^{\textrm{post-mrg}}(t)\, \theta\left(t-t_{\textrm{match}}^{22}\right)\,.
\end{align}

@@

As described above, an EOB waveform with dynamical tides depends on 11 intrinsic parameters: the mass ratio $0\leq q=M_B/M_A\leq1$, the dimensionless components of the spins $\boldsymbol{S}_{A,B}$ along the Newtonian angular momentum ${\boldsymbol{L}}_N$, $-1 \leq \chi_{A,B}=(\boldsymbol{S}_{A,B}\cdot \hat{\boldsymbol{L}}_N)/M_{A,B}^2 \leq1$, the adiabatic quadrupolar tidal polarizabilities $\Lambda_2^{A,B}$, the adiabatic octupolar tidal polarizabilities $\Lambda_3^{A,B}$, the $\ell=2,3$ $f$-mode angular frequencies $\omega_{0\ell}^{A,B}$. To reduce the dimensionality of the intrinsic parameter space, we exploit universal (i.e., equation-of-state-independent) relations that are available in the literature. In particular, Eq.~(60) of Ref.~\cite{Yagi:2013sva} with the coefficients from the first row of Table I therein expresses $\Lambda_3^{A,B}$ as a function of $\Lambda_2^{A,B}$ with an accuracy to within about 10\%, depending on the equation of state and the value of $\Lambda_2^{A,B}$. While this relation has been fitted to NR data only for $1 \leq \Lambda_2^{A,B}\leq10^6$, we use it also outside such range. However, this formula is divergent at small values of $\Lambda_2^{A,B}$, thus in such regime we replace it with a polynomial function that vanishes at 0, equals the universal relation at $\Lambda_2^{A,B}=10^{-2}$, and best-fits it for $10^{-5}\leq\Lambda_2^{A,B}\leq10^{-2}$ Explicitly, putting $\zeta_{A,B}=\frac{\log\Lambda^{A,B}_2}{10}$,
\begin{widetext}
\begin{equation}
\Lambda_3^{A,B}=\left\{\begin{array}{ll}
\Lambda_2^{A,B}\left[0.441 - 34.6\,\Lambda_2^{A,B} + 1760(\Lambda_2^{A,B})^2\right]\,,&\textrm{if}\quad0\leq\Lambda_2^{A,B}\leq10^{-2}\\
&\\
\exp{\left[-1.15 +11.8\,\zeta_{A,B}+2.51\,\zeta_{A,B}^2-1.31\,\zeta_{A,B}^3+0.252\,\zeta_{A,B}^4\right]}\,,&\textrm{otherwise}\,.
\end{array}\right.
\end{equation}
\end{widetext}
Eq. (3.5) of Ref.~\cite{Chan:2014kua} with the coefficients from the first column of Table I therein gives $\omega_{02}^{A,B}$ as a function of $\Lambda_2^{A,B}$ to within a few percent, while the same equation but with the coefficients from the second column of Table I therein gives $\omega_{03}^{A,B}$ as a function of $\Lambda_3^{A,B}$ to within a few percent. Putting $\xi_{A,B}=\log\Lambda_2^{A,B}$, over the fitting range used by Ref.~\cite{Chan:2014kua} (i.e., $0\leq \xi_{A,B} \leq 9$), we set
\begin{align}
M_{A,B}\omega_{02}^{A,B}&=
0.182 -6.836\times10^{-3}\xi_{A,B}\nonumber\\
&-4.196\times10^{-3}\xi_{A,B}^2+5.215\times10^{-4}\xi_{A,B}^3\nonumber\\
&-1.857\times10^{-5}\xi_{A,B}^4\,;\label{URomega2}
\end{align}
outside this range, we set $M_{A,B}\omega_{02}^{A,B}=0.182$ if $\xi_{A,B} < 1$ and $M_{A,B}\omega_{02}^{A,B}=0.161$ if $\xi_{A,B} >9$, that is Eq.~(\ref{URomega2}) evaluated for $\xi_{A,B}=1$ and $\xi_{A,B}=9$, respectively. Moreover, putting $\Upsilon_{A,B}=\log\Lambda_3^{A,B}$, over the fitting range used by Ref.~\cite{Chan:2014kua} (i.e., $-1 \leq \Upsilon_{A,B} \leq 10$), we set
\begin{align}
M_{A,B}\omega_{03}^{A,B}&=
\,0.2245 -1.5\times10^{-2}\Upsilon_{A,B}\nonumber\\
&-1.412\times10^{-3}\Upsilon_{A,B}^2+1.832\times10^{-4}\Upsilon_{A,B}^3\nonumber\\
&-5.561\times10^{-6}\Upsilon_{A,B}^4\,;\label{URomega3}
\end{align}
outside this range, we set $M_{A,B}\omega_{03}^{A,B}=0.2379$ if $\Upsilon_{A,B} < -1$ and $M_{A,B}\omega_{03}^{A,B}=0.1165$ if $\Upsilon_{A,B} >10$, that is Eq.~(\ref{URomega3}) evaluated for $\Upsilon_{A,B}=-1$ and $\Upsilon_{A,B}=10$, respectively.
Thanks to these relations, the waveform only depends on 5 intrinsic parameters, which we denote as $\boldsymbol{x}=\{q, \chi_A,\chi_B, \Lambda_2^A,\Lambda_2^B\}$. We note that this list does not include the total mass $M$ of the system. Of course, the point-mass part of the model is scale-invariant. As to the inspiral-plunge phase, in this model tidal corrections only depend on $X_A=1/(1+q)$, $X_B=q/(1+q)$, and $\Lambda_2^{A,B}$, which is inherent to the post-Newtonian analysis of tidal effects~\cite{Flanagan:2007ix}. Finally, the simple model of the post-merger signal that we employ does not depend on $M$ either, although a more sophisticated description will have to include such dependence to account for the phenomenology that is observed in NR simulations of this phase.
